Homeostatic mechanisms may shape the type and duration of oscillatory modulation.

### 2019-04-22 main.tex
Updates from Overleaf

%% Useful packages
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{graphicx}
\usepackage[colorlinks=true, allcolors=blue]{hyperref}
\usepackage{setspace}
\usepackage{lineno}  % For reviewers
\usepackage{authblk}
\usepackage{blindtext}

@@

% Make a pretty abstract box...
% \renewenvironment{abstract}
% {\begin{quote}
% \noindent \rule{\linewidth}{.5pt}\par{\bfseries \abstractname.}}
% {\medskip\noindent \rule{\linewidth}{.5pt}
% \end{quote}
% }

@@

\begin{document}
\maketitle
% \linenumbers
% -----------------------------------------------------
% -----------------------------------------------------
% -----------------------------------------------------
\begin{abstract}
Neural oscillations are observed ubiquitously in the mammalian brain. However the stability of oscillations is highly variable. Some oscillations are \textit{tonic}, lasting for seconds or even minutes; others are unstable, appearing only as a single-cycle \textit{burst}. In a model of hippocampal neurons, we use numerical simulations to show how these different forms of rhythm stability can interact with activity-dependent homeostasis to profoundly alter the modulatory effect of neural oscillations. Under homeostasis, tonic oscillations that are synaptically excitatory have a paradoxical effect; they decrease excitability and desynchronizing firing. Tonic oscillations that are synaptically inhibitory--like those in a real hippocampus--fail to generate new action potentials and so provoke no homeostatic response. This may explain why the theta rhythm in hippocampus is synaptically inhibitory: inhibitory oscillations don't raise the firing threshold, as excitatory oscillations do, and so can preserve each cell's dynamic range. Based on these simulations, we also speculate that homeostasis may explain why excitatory intra-cortical and intra-layer oscillations often appear as bursts. In our model bursts minimally interact with the slow homeostasis time constant and so retain typical excitatory effects.\\
\end{abstract}

% -----------------------------------------------------
% -----------------------------------------------------
% -----------------------------------------------------
\section*{Introduction}
\lettrine[loversize=0,nindent=0,realheight=true]{N}{euromodulation\ and\ homeostasis} are inter-linked but opposing phenomena. Modulation perturbs excitability, which we define as the propensity for a stimulus to elicit an action potential. Homeostasis acts to this quench perturbation, driving the excitability of the cell back to a biologically desirable set point \cite{LeMasson1993,Abbott1993}. While the interplay between chemical modulators and homeostasis has been studied for more than 20 years \cite{LeMasson1993,Abbott1993,Golowasch1999,Marder2014,Marder2015,Gutierrez2013}, the relationship between network-level synaptic modulations--like neural oscillations--and homeostasis is not well understood, on either theoretical or empirical grounds. Like neuromodulators, neural oscillations alter excitability and firing statistics. Uniquely, though, oscillations create synchronous windows of activity \cite{Lisman2013,Voytek2015}. Temporally grouping action potentials into windows improves signal to noise and increases the number of coincident firing events \cite{Chen2013,Zhou2015,Voytek2015a,Peterson2017}, driving learning at individual synapses \cite{Muller2011,Song2000,Markram1997}.

>>>>

%% Useful packages
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{graphicx}
\usepackage[colorlinks=true, allcolors=blue]{hyperref}
\usepackage{setspace}
\usepackage{lineno}  % For reviewers
\usepackage{authblk}
\usepackage{blindtext}
\usepackage{fancyhdr}

@@

\pagestyle{fancy}
\rhead{Homeostasis and oscillations}

@@

\begin{document}
\maketitle
\linenumbers
% -----------------------------------------------------
% -----------------------------------------------------
% -----------------------------------------------------
\begin{abstract}
Neural oscillations are observed ubiquitously in the mammalian brain. However the stability of oscillations is highly variable. Some oscillations are \textit{tonic}, lasting for seconds or even minutes; others are unstable, appearing only as a single-cycle \textit{burst}. In a model of hippocampal neurons, we use numerical simulations to show how these different forms of rhythm stability can interact with activity-dependent homeostasis to profoundly alter the modulatory effect of neural oscillations. Under homeostasis, tonic oscillations that are synaptically excitatory have a paradoxical effect; they decrease excitability and desynchronizing firing. Tonic oscillations that are synaptically inhibitory--like those in a real hippocampus--fail to generate new action potentials and so provoke no homeostatic response. This may explain why the theta rhythm in hippocampus is synaptically inhibitory: inhibitory oscillations don't raise the firing threshold, as excitatory oscillations do, and so can preserve each cell's dynamic range. Based on these simulations, we also speculate that homeostasis may explain why excitatory intra-cortical and intra-layer oscillations often appear as bursts. In our model bursts minimally interact with the slow homeostasis time constant and so retain typical excitatory effects.\\
\end{abstract}

% -----------------------------------------------------
% -----------------------------------------------------
% -----------------------------------------------------
\noindent
\textsc{\textbf{New and Noteworthy}} In a model of hippocampal neurons we report a paradoxical result: Ca\textsuperscript{2+}-mediated homeostasis causes normally synchronizing AMPAergic oscillations to become inhibitory and desynchronizing. We conjecture this could be why the theta rhythm is inhibitory. In pyramidal cells GABA doesn't generate new action potentials and so provokes no homeostatic response. The intricate interplay of neuromodulators, like acetylcholine, with homeostasis is well known. The interplay between oscillations and homeostasis might be as complex, and as important.

\section*{Introduction}
\lettrine[loversize=0,nindent=0,realheight=true]{N}{} euromodulation and homeostasis are inter-linked but opposing phenomena. Modulation perturbs excitability, which we define as the propensity for a stimulus to elicit an action potential. Homeostasis acts to this quench perturbation, driving the excitability of the cell back to a biologically desirable set point \cite{LeMasson1993,Abbott1993}. While the interplay between chemical modulators and homeostasis has been studied for more than 20 years \cite{LeMasson1993,Abbott1993,Golowasch1999,Marder2014,Marder2015,Gutierrez2013}, the relationship between network-level synaptic modulations--like neural oscillations--and homeostasis is not well understood, on either theoretical or empirical grounds. Like neuromodulators, neural oscillations alter excitability and firing statistics. Uniquely, though, oscillations create synchronous windows of activity \cite{Lisman2013,Voytek2015}. Temporally grouping action potentials into windows improves signal to noise and increases the number of coincident firing events \cite{Chen2013,Zhou2015,Voytek2015a,Peterson2017}, driving learning at individual synapses \cite{Muller2011,Song2000,Markram1997}.
